### 2020-12-15 paper_15_12_2020/letter/letter.tex

\documentclass[version=last,12pt,{"/home/prdm0/R/x86_64-pc-linux-gnu-library/4.0/komaletter/rmarkdown/templates/pdf/resources/maintainersDelight"},letterpaper,]{scrlttr2}

% For Debuging the Layout  ----------------------------------------------------
% \LoadLetterOption{visualize}
% \showfields{head,address,location,refline,foot}
% \setkomafont{field}{\color{blue}}
% -----------------------------------------------------------------------------

\usepackage{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
% \usepackage[margin=1in]{geometry} interfers with the areaset in my lco_default
%\usepackage[margin=1in]{geometry}
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdfauthor={Prof.~Dr.~Pedro Rafael D. Marinho},
            colorlinks=true,
            linkcolor=red,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
% Make links footnotes instead of hotlinks:
\renewcommand{\href}[2]{#2\footnote{\url{#1}}}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\makeatletter
\usepackage{graphicx}
% grffile has become a legacy package: https://ctan.org/pkg/grffile
\IfFileExists{grffile.sty}{%
\usepackage{grffile}
}{}
% set default figure placement to htbp
\def\fps@figure{htbp}
\setkomavar{signature}{Prof.~Dr.~Pedro Rafael D. Marinho\\
(DSc in Statistics from Federal University of Pernambuco)}
\makeatother

\KOMAoptions{fromemail=true}
\setkomavar{fromemail}{pedro.rafael.marinho@gmail.com}
\KOMAoptions{fromurl=true}
\setkomavar{fromurl}{https://prdm0.rbind.io/}

\setkomavar{yourmail}{pedro.rafael.marinho@gmail.com}

\setkomavar{subject}{A new family of distributions: properties and
applications}
\setkomavar{fromname}{Prof.~Dr.~Pedro Rafael D. Marinho}
\setkomavar{fromaddress}{Department of Statistics\\Federal University of
Paraíba\\João Pessoa, Paraíba - PB, Brazil}

\begin{document}

\begin{letter}{\textbf{Editor-in-Chief: Enrico
Colosimo}\\\textbf{Brazilian Journal of Probability and Statistics -
BJPS}\\}
\opening{Dear \textbf{BJPS} Editor,}

we would like to present the paper entitled \textbf{A new family of
distributions: properties and applications} to the BJPS. This article
introduces a new family by combining the Marshall and Olkin-G and
Gamma-G classes. The family has only two extra shape parameters and can
be a better model than other existing classes of distributions.
Simulations are performed to verify the consistency of the estimators.
Its flexibility is shown using two real data sets.

This article submitted to this journal is co-authored by researchers
Gauss M. Cordeiro (PhD in Statistics from Imperial College, 1982) of the
Statistics Department of Federal University of Pernambuco and Dra. Maria
do Carmo S. Lima Federal of the Statistics Department of Federal
University of Pernambuco.

This material has not been submitted to any other Journal.

\closing{Kind regards,}

\end{letter}

\end{document}