### 2022-05-17 paper/reviews/answers/letter_answers.tex

\documentclass[version=last,12pt,{"/home/prdm0/R/x86_64-pc-linux-gnu-library/4.2/komaletter/rmarkdown/templates/pdf/resources/maintainersDelight"},letterpaper,]{scrlttr2}

% For Debuging the Layout  ----------------------------------------------------
% \LoadLetterOption{visualize}
% \showfields{head,address,location,refline,foot}
% \setkomafont{field}{\color{blue}}
% -----------------------------------------------------------------------------

\usepackage{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
% \usepackage[margin=1in]{geometry} interfers with the areaset in my lco_default
%\usepackage[margin=1in]{geometry}
\usepackage[unicode=true,pdfversion=1.6]{hyperref} % PrintScaling needs 1.6
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdfauthor={Prof.~Dr.~Pedro Rafael D. Marinho},
            colorlinks=true,
            linkcolor=red,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true,
            pdfprintscaling=None} % signal the PDF viewer "no auto scale", PDF >= 1.6
\urlstyle{same}  % don't use monospace font for urls
% make links footnotes instead of hotlinks (default)
\renewcommand{\href}[2]{#2\footnote{\url{#1}}}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\makeatletter
\usepackage{graphicx}
% grffile has become a legacy package: https://ctan.org/pkg/grffile
\IfFileExists{grffile.sty}{%
\usepackage{grffile}
}{}
% set default figure placement to htbp
\def\fps@figure{htbp}
\setkomavar{signature}{Prof.~Dr.~Pedro Rafael D. Marinho\\
(DSc in Statistics from Federal University of Pernambuco)}
\makeatother

\KOMAoptions{fromemail=true}
\setkomavar{fromemail}{pedro.rafael.marinho@gmail.com}
\KOMAoptions{fromurl=true}
\setkomavar{fromurl}{https://prdm0.rbind.io/}

\setkomavar{yourmail}{pedro.rafael.marinho@gmail.com}

\setkomavar{subject}{Marshall and Olkin-G and Gamma-G family of
distribution: properties and applications}
\setkomavar{fromname}{Prof.~Dr.~Pedro Rafael D. Marinho}
\setkomavar{fromaddress}{Department of Statistics\\Federal University of
Paraíba\\João Pessoa, Paraíba - PB, Brazil}

\begin{document}

\begin{letter}{\textbf{Replies to reviewers}\\\textbf{REVSTAT}\\}
\opening{Dear \textbf{REVSTAT} Executive Editor and Reviews,}

in this letter, I will respond to each of the reviewers valuable
contributions. We try to accept all suggestions, aiming at a significant
improvement of our article. After each questioning, the respective
answers will be given. As suggested by the second reviewer, the title of
the paper was modified and better represents the proposal of the work.

\textbf{REVIEW I}:

1 -- The equality \(h(x)=f(x)/(1-F(x))\) is a well known relation
between hazard, density and survival functions. It is not a definition
of density or risk functions. Therefore, in page 3, line 12, the part of
the sentence ``defined by \(h(x)=f(x)/(1-F(x))\)'' should be omitted.

\emph{Response}:

2 -- The expression (3.1) in page 6 corresponds to the log-likelihood of
a MO-\(\Gamma\)-\(G\) distribution, considering a Weibull distribution
for \(G\) (although there is an error in the last sum of line 1), not a
general distribution with a \(\eta\) vector of parameters. Therefore,
the text or the expression should be written accordingly.

\emph{Response}:

3 -- It is not clear why it is considered the Weibull distribution (for
\(G\)) in Figure 1, eventually in expression (3.1) and in the
applications, whereas for the simulations it is chosen the Dagum
distribution. These choices should be clarified.

\emph{Response}:
\textcolor{red}{Indeed, the way in which the application section was written suggests that an attempt was made to fix the $G\sim Weibull$ distribution as one of the necessary assumptions for the analysis. What actually happened is that for the datasets used, the Weibull distribution appeared to be a suitable distribution.}

\textcolor{red}{In order to remove some ambiguity, we opted to extend the simulation studies, where we also include the simulations, considering $G\sim Weibull$.
Indeed, the way in which the application section was written suggests that an attempt was made to fix the $G\sim Weibull$ distribution as one of the necessary assumptions for the analysis. What actually happened is that for the datasets used, the Weibull distribution appeared to be a suitable distribution.}

\textcolor{red}{In order to remove some ambiguity, we opted to extend the simulation studies, where we also include the simulations, considering $G\sim Weibull$.}

4 -- As the Dagum distribution is used in the simulations, the generated
PDF should also be in Table 1.

\emph{Response}:
\textcolor{red}{Thanks for the observation. The Dagum distribution was included in Table 1.}

5 -- As the \texttt{goodness.fit} function gives several statistics, the
authors should justify why they only chose the Anderson Darling and
Cramer-von Mises statistics.

\emph{Response}:
\textcolor{red}{The use of the $A^*$ and $W^*$ statistics, Anderson-Darling and Cramer-von Mises statistics, respectively, corrected by Chen and Balakrishnan, in the paper entitled "A General Purpose Approximate Goodness-of-Fit Test" (see [link](https://www.tandfonline.com/doi/abs/10.1080/00224065.1995.11979578)) are suitable for comparing models that are not necessarily embedded. Therefore, considering the statistics returned by the `goodness.fit` function and considering the distributions compared, we understand that the statistics $A^*$ and $W^*$ are the most suitable for making the comparisons.}

\textcolor{red}{The **AdequacyModel** package was developed by one of the authors of this paper and was published at [https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0221487](https://journals .plos.org/plosone/article?id=10.1371/journal.pone.0221487). A code inspection of the `goodness.fit` function can be done by looking at the [link](https://github.com/prdm0/AdequacyModel/blob/master/R/goodness.fit.R) on GitHub, where it is hosted the development versioning of the package.}

\emph{Minor points:}

1 -- page 1, line 8 of Section 1, correct for ``vector''.

\emph{Response}:

2 -- page 2: confirm the expressions of \(Q_{MO-G}(u)\) (line 9) and
\(W\) (line22).

\emph{Response}:

3 -- page 2, line -4 of Section 1, correct for ``are''.

\emph{Response}:

4 -- page 3, line 4, where it is written ``gamma density unit scale and
shape \(a>0\),'' I suggest replacing it by ``gamma density with unit
scale parameter and shape parameter \(a>0\),''.

\emph{Response}:

5 -- page 3, \(3^{rd}\) paragraph, replace ``Table 2'' by ``Table 1''.

\emph{Response}:

6 -- page 3, line-8, correct for ``(0,1)''.

\emph{Response}:

7 -- Figure 1: I suggest to change the lines to different types (even
letting the colours as they are) so as to enable the interpretation in a
black and white print. Also, the letters assigned to the parameters must
agree with those in Table 1 (and, eventually, with those in expression
(3.1)).

\emph{Response}:
\textcolor{red}{Thanks a lot for the suggestion. All changes have been made. In fact, in a black and white print, solid lines will not help to differentiate the graphics.}

8 -- page 3, lines-8,-7, where it is written ``of the baseline \(Q_F\)
\(Q_G(.)\).'' I suggest replacing it by ``of the \(Q_F\) of the baseline
distribution \(G\), \(Q_G(.).\)''.

\emph{Response}:

9 -- page 3, line-6, correct for ``(1.1) and''.

\emph{Response}:

10 -- page 7: the label of the table should be on the top, like the rest
of the tables.

\emph{Response}:

11 -- page 11, line 4, correct for ``distribution''.

\emph{Response}:

12 -- In expressions (6.4) and (6.6), try to improve the output as the
reference numbers are to close to the formulas.

\emph{Response}:

13 -- I suggest to make the labels of Figure 1 and of tables in \LaTeX,
not in R.

\emph{Response}:
\textcolor{red}{Thanks for the sugestion. The change suggestion has been implemented.}

14 -- The references ``R Core Team. \ldots{}'' and ``Rizzo, \ldots{}''
are in the wrong order.

\emph{Response}:

\textbf{REVIEW II}:

1 - The article is written in an overly summarized form and is sometimes
unattractive to the reader of any journal. That alone could lead to your
rejection. In addition, I do suggest reviewing the use of English
carefully, and considerable rewording and pruning to make the paper more
concise and precise.

\emph{Response}:

2 - When I see a new paper about a new distribution, I wonder if it is
due to any interesting properties or it arises naturally in some
observable process like natural phenomena. I would like to see these
arguments defended more clearly in the article by the authors. Could you
do that?

\emph{Response}:

3 - The bibliography does not have many recent publications such as
Martinez-Florez et al. (2020) and Magalhaes et al.~(2020). I also
suggest authors to create the state-of-art related to this topic in the
first section that definitively needs to be improved.

\emph{Response}:

4 - I expect to see more detailed discussion in the simulation study and
real data analysis. Could you revise these sections, including the
conclusion section that is actually weak?

5 - Suggestions: (i) Rewrite the paper title since the current one is
nothing enlightening. How about naming the proposed distribution?. (ii)
Place section 6 (properties) before section 4 (simulation). (iii)
Reorganize the presentation of the tables. (iv) Put the simulation code
in Supplementary material.

\emph{Response}:

\closing{Kind regards,}

\end{letter}

\end{document}